Commutative Character Sheaves on smooth group schemes over finite fields

### 2016-01-13 followup.tex
Rewrote the section on smooth models as a step toward geometrization of types.

As proved in \cite{kim:07a}, all supercuspidal representations of unramified $p$-adic groups can be built from the data of \cite{yu:01a}*{\S 15}.
The type, in the sense of \cite{bushnell-kutzko:98a}, of the resulting supercuspidal representations can be constructed directly from Yu's data. 
Let us refer to the types arising in this manner as {\it Yu types}.
%In this section we show that every Yu type is geometrizable.
% in a sense which will be made clear below.
%
A Yu type is then a pair $(\,^\circ K,\,^\circ \rho)$, where $\,^\circ K$ is a compact group and $\,^\circ \rho$ is a representation of $\,^\circ K$, constructed in a certain way. 
%
Regarding the construction of the type $(\,^\circ  K,\,^\circ \rho)$, the paper \cite{yu:models} concludes with the following provocative observation.

\begin{quotation}
{\it Therefore, up to some linear characters, all the ingredient representations are on groups of
 the form $\underline{H}(\mathcal{O})$, where $\underline{H}$ is a smooth group scheme over $\mathcal{O}$, and the representations are inflated from $\underline{H}(\kappa)$. These results suggest that algebraic geometry and group schemes should play an
important role in the representation theory of $p$-adic groups.} \cite{yu:03a}
\end{quotation}

In Section~\ref{sec:types} we use Theorem~\ref{thm:geo} to results push this observation further by geometrizing the representation $\,^\circ \rho$ of $\,^\circ K$: we find a group scheme $G$ over $\Fq$ such that $G(\Fq) = \,^\circ K$ and a virtual Weil sheaf $\mathcal{F}$ on $G$ such that $\trFrob{\cs{F}} = \trace(\,^\circ \rho)$.
%from which the character of $\,^\circ \rho$ can be recovered using the trace of Frobenius.

@@

By \cite{yu:models}*{Prop 10.2} there is a sequence of morphisms of affine smooth group schemes of finite type over $\OK$
\[
\underline{G}^0 \to \underline{G}^1 \to \cdots \to \underline{G}^d = \underline{G}
\]
such that $\underline{G}^i(\OK) = \,^\circ K^i$ for $i=0,\ldots d$.
Also, for each $i=0,\ldots d$, there is morphism of affine smooth group schemes of finite type over $\OK$ 
\[
\underline{J}^i \to \underline{G}
\] 
such that $\underline{J}^i(\OK) = J^i$ as a subgroup of $\,^\circ K$.

%
As explained in \cite{yu:03a}*{\S 10.4}, the group scheme $\underline{G}$ comes with the following auxilliary data:
\begin{enumerate}
\item[(Ji)]
	affine smooth group schemes $\underline{J}^i$, for $i=0, 1, \ldots, d$, together with morphisms $\underline{J}^i \to \underline{G}$ such that \[\underline{J}^i(\OK) = J^i,\]
\item[(Gj)] affine smooth group schemes $\underline{G}^{j}$, for $j=0, \ldots, d$, together with morphisms $\underline{G}^{j}\to \underline{G}$ such that 
\[
\underline{G}^{j}(\OK) = J^0\cdots J^i.
\] 
\end{enumerate}

@@

With these group schemes, the other groups and group homomorphisms appearing in Section~\ref{ssec:review} may be put into the framework needed for geometrization.

@@

Let us return to the list of Section~\ref{ssec:review} and consider how we have recast each part.
\begin{enumerate}
\item[(0)] The compact group $\,^\circ K$ is replaced by the smooth group scheme $\underline{G}$ over $\OK$.
\item[(1)] The compact open subgroups $J^i \hookrightarrow \,^\circ K$ are replaced by the morphisms $\underline{J}^i \to \underline{G}$ of smooth group schemes over $\OK$.
\item[(2)] The representation $\varpi_0$ is obtained by pullback from a representation $\sigma_0$ of $\underline{J}^{0,\reductive}_\Fq$, which makes use of the following morphisms of group schemes.
\[
\underline{J}^{0} \longleftarrow \underline{J}^{0}_\Fq \to \underline{J}^{0,\reductive}_\Fq.
\]
\item[(3)] The quasicharacters $\varphi_i : \underline{J}^i(\OK) \to \CC^\times$ have not yet been geometrized.
\item[(4)] The quotient $J^{i+1}\to V_{i+1}$ has been replaced with the following morphisms of group schemes.
\[
\underline{J}^{i+1} \longleftarrow \underline{J}^{i+1}_\Fq \to \underline{J}^{i+1,\reductive}_\Fq \to V_{i+1,\Fq}.
\]
\item[(5)] The symplectic pairing $\langle\ ,\ \rangle_{i+1} : V_{i+1}\times V_{i+1} \to C_{i+1}$ is given by 
\item[(6)]
\item[(7)] The group homomorphism $\,^\circ f_i : J^0 \cdots J^i \to \Sp(V_{i+1})$ has been replaced by the following morphisms of group schemes.
\[
\underline{G}^{i} \longleftarrow \underline{G}^{i}_\Fq \to \underline{G}^{i,\reductive}_\Fq \to \Sp(V_{i,\Fq}),
\]
and $\Sp(V_{i,\Fq})$ is itself a reductive algebraic group over $\Fq$.
\item[(8)] The quotient $J^{i+1} \to H^{i+1}$ has been replaced by the following morphisms of group schemes.
\[
\underline{J}^{i+1} \longleftarrow \underline{J}^{i+1}_\Fq \to H^{i+1}_\Fq.
\] 
\item[(9)]
The Heisenberg group $V_{i+1}^\sharp$ is replaced by a unipotent algebraic group $V_{i+1,\Fq}^\sharp$ over $\Fq$, and the group isomorphism $H^{i+1} \iso V_{i+1}^\sharp$ is replaced by an isomorphism $H^{i+1}_\Fq \iso V_{i+1,\Fq}^\sharp$ of algebraic groups over $\Fq$.
\item[(10)]

\end{enumerate}

\subsection{Geometrization of the character of the Heisenberg-Weil representation}

@@

>>>>

As proved in \cite{kim:07a}, all supercuspidal representations of unramified $p$-adic groups can be built from the data of \cite{yu:01a}*{\S 15}.
The type, in the sense of \cite{bushnell-kutzko:98a}, of the resulting supercuspidal representations can be constructed directly from Yu's data. 
Let us refer to the types arising in this manner as {\it Yu types}.
%In this section we show that every Yu type is geometrizable.
% in a sense which will be made clear below.
%
A Yu type is then a pair $(\,^\circ K,\,^\circ \rho)$, where $\,^\circ K$ is a compact group and $\,^\circ \rho$ is a representation of $\,^\circ K$, constructed in a certain way. 
%
In this section we use Theorem~\ref{thm:geo} to geometrize the representation $\,^\circ \rho$ of $\,^\circ K$: we find a group scheme $G$ over $\Fq$ such that $G(\Fq) = \,^\circ K$ and a virtual Weil sheaf $\mathcal{F}$ on $G$ such that $\trFrob{\cs{F}} = \trace(\,^\circ \rho)$.
%from which the character of $\,^\circ \rho$ can be recovered using the trace of Frobenius.

@@

\begin{enumerate}
\item[(K)]
By \cite{yu:models}*{Prop 10.2} there is a sequence  
\[
\underline{G}^0 \to \underline{G}^1 \to \cdots \to \underline{G}^d = \underline{G}
\]
of morphisms of affine smooth group schemes of finite type over $\OK$ such that, on $\OK$-points it gives the sequence $\,^\circ K^0 \subseteq \,^\circ K^1 \subseteq \cdots \subseteq \,^\circ K^d= \,^\circ K$ above.
\item[(J)]
As explained in \cite{yu:03a}*{\S 10.4}, for each $i=0,\ldots d$, there is morphism of affine smooth group schemes of finite type over $\OK$ 
\[
\underline{J}^i \to \underline{G}
\] 
such that $\underline{J}^i(\OK) = J^i$ as a subgroup of $\,^\circ K$ and such that the image of the $\OK$-points under the multiplication map $\underline{J}^0 \times \cdots \times \underline{J}^i \to \underline{G}$ is $\,^\circ K^i$, for $i=0, \ldots , d$.
\item[(S)]
There is a natural action of $\underline{G}^i$ on $\underline{J}^{i+1}$ in the category of smooth affine group schemes over $\OK$ so that the group
\[
\underline{G}^i \ltimes \underline{J}^{i+1}
\]
has $\OK$-points
\[
(\underline{G}^i \ltimes \underline{J}^{i+1})(\OK) = \,^\circ K^i \ltimes J^{i+1}.
\]
\end{enumerate}

@@

@@

With all this, we may revisit the quotients appearing in Section~\ref{ssec:review}:
\[
\begin{tikzcd}
1 \arrow{r} & \underline{J}^{i+1}  \arrow{r} & \underline{G}^i \ltimes \underline{J}^{i+1} \arrow{r} & \underline{G}^i \arrow{r}  & 1\\
1 \arrow{r} & \underline{J}^{i+1}_\Fq \arrow{d} \arrow{u} \arrow{r} & \underline{G}_\Fq^i \ltimes \underline{J}_\Fq^{i+1} \arrow{d} \arrow{u} \arrow{r} & \underline{G}_\Fq^i \arrow{r} \arrow{d} \arrow{u} & 1\\ 
1 \arrow{r} & V_{i+1}^\sharp \arrow{r} & \Sp(V_{i+1}) \ltimes V_{i+1}^\sharp \arrow{r} & \Sp(V_{i+1}) \arrow{r} & 1,
\end{tikzcd}
\]
where the last two rows are now understood as forming a diagram in the category of algebraic groups over $\Fq$. 
This realizes the Jacobi group $\Sp(V_{i+1}) \ltimes V_{i+1}^\sharp$ as a quotient of the special fibre of the smooth group scheme $\underline{G}^i \ltimes \underline{J}^{i+1}$ over $\OK$.

We may now revisit the ingredients in the construction of the representation $\,^\circ \rho$ of $\underline{G}(\OK)$ along the lines indicated by Yu and recalled at the beginning of this section.
\begin{enumerate}
\item The continuous representation $\varrho$ of $\,^\circ K^0$ is a representation of $\underline{G}^0(\OK)$ obtained by inflation along $\underline{G}^0(\OK) \to \underline{G}^0(\Fq)$. 
Let $\sigma$ be that representation of $\underline{G}^0(\Fq) = \underline{G}^0_\Fq(\Fq)$.
In fact, $\sigma$ is itself obtained by pulling back a representation $\sigma^0$ along the $\Fq$-points of the quotient $\underline{G}^0_\Fq \to (\underline{G}^0)_\Fq^\reductive$.
\item Quasicharacters $\varphi_i : \underline{G}^i(\OK) \to \CC^\times$, for $i=0, \ldots d$.
\item The representation $h_{i+1}^*(W_{i+1})$ of $(\underline{G}^i \ltimes \underline{J}^{i+1})(\OK)$ is obtained by pulling back a representation along 
\[
(\underline{G}^i \ltimes \underline{J}^{i+1})(\OK) \to (\underline{G}^i \ltimes \underline{J}^{i+1})(\Fq).
\]
Let $w_{i+1}$ be that representation of $(\underline{G}^i \ltimes \underline{J}^{i+1})(\Fq) = (\underline{G}_\Fq^i \ltimes \underline{J}^{i+1}_\Fq)(\Fq)$. 
Then $w_{i+1}$ is itself obtained by pulling back the representation $W_{i+1}$ along the $\Fq$-points of the quotient
\[
\underline{G}_\Fq^i \ltimes \underline{J}^{i+1}_\Fq \to 
\Sp(V_{i+1}) \ltimes V_{i+1}^\sharp.
\]
\end{enumerate}
Regarding the construction of the type $(\,^\circ  K,\,^\circ \rho)$, the paper \cite{yu:models} concludes with the following provocative observation, which we have just revisited.
\begin{quotation}
{\it Therefore, up to some linear characters, all the ingredient representations are on groups of
 the form $\underline{H}(\mathcal{O})$, where $\underline{H}$ is a smooth group scheme over $\mathcal{O}$, and the representations are inflated from $\underline{H}(\kappa)$. These results suggest that algebraic geometry and group schemes should play an
important role in the representation theory of $p$-adic groups.} \cite{yu:03a}
\end{quotation}

\subsection{Geometrization of characters of the ingredient representations}

\todo{Everything below here is under construction}
Recall that $\sigma^0$ is a representation of $(\underline{G}^0)_\Fq^\reductive$.
Suppose $\pi_0((\underline{G}^0)_\Fq^\reductive)$ is cyclic.
Let $A^0$ be a Weil sheaf on $(\underline{G}^0)_\Fq^\reductive$ such that, as a sheaf on $(\underline{G}^0)_\Fq^\reductive \otimes \bFq$, it is a character sheaf.
The existence of $A^0$ follows from \cite{Lusztig:CSDG}*{}.
Let $A$ be the Weil sheaf on $\underline{G}^0_\Fq$ obtained by pullback along $\underline{G}^0_\Fq \to (\underline{G}^0)_\Fq^\reductive$.

Let $n$ be an integer greater than the supremum of the depths of $\varphi_i : \underline{G}^i(\OK)\to \CC^\times$, for $i=0,\ldots, d$.
Then, for $i=0,\ldots, d$ we have the commuting diagram 
\[
\begin{tikzcd}
\end{tikzcd}
\]

@@

\begin{lemma}
Let $\underline{G}$ be a smooth group scheme over $\OK$.
Let $\varphi : \underline{G}(\OK) \to \CC^\times$ be a quasicharacter/continuous linear character. 
The there is a Weil sheaf $\cs{F}$ on $\Gr(\underline{G})$ such that $\trFrob{\cs{F}} = \trace \varphi$.
\end{lemma}
